Modeling site-specific amino-acid preferences deepens phylogenetic estimates of the divergence of viral proteins.

### 2018-03-15 manuscript/manuscript.tex
added x-axis label to empirical trees plot
began to reorganize supp files

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figures/empirical_trees.pdf}}
\caption{\label{fig:empirical_trees}
\textbf{Effect of site-specific amino-acid preferences and $\Gamma\omega$ rate variation on HA branch length estimation.} 
The branch lengths of the HA tree are optimized using the indicated ExpCM or GY94 model. 
The amino-acid preferences defining the model (ExpCM) or implied by the model (GY94) are shown as logo plots for 15 sites in HA; the full set of experimentally measured amino-acid preferences are in \ref{suppfig:prefs_doud}, \ref{suppfig:prefs_lee}, and \ref{suppfig:prefs_average}. 
The ExpCM's use amino-acid preferences measured in deep mutational scanning of an H1 HA~\citep{doud2016accurate}, an H3 HA~\citep{lee2018deep}, or the average of the measurements for these two HAs.
Circle denotes the H1 clade and triangle denotes the H3 clade.
\jdbcomment{Add a x-axis label (probably can just be for bottom-most plots) giving the units: codon substitutions per site.} 
\jdbcomment{Add a line to this legend about how trees are rooted, such as: ``Each tree is separately midpoint rooted''---or if that is not what was done, then add something that indicates how it was done. If it isn't what was done, can you remind me of rationale for not doing that?}
}
\end{figure}

@@

@@

\begin{suppfile}
\caption{
\label{suppfile:WSN_Perth_map}
List of alignable sites between H1 HA and H3 HA. 
This file provides a conversion between the numbering scheme we use in the paper (sequential numbering of just the alignable sites) to sequential numbering of the H1 HA reference sequence A/Wilson Smith/1933 and the H3 HA reference sequence A/Perth/2009. 
The sequential numbering for the HA strains starts with ``site 1" as the second codon (first non-methionine codon). 
\jdbcomment{The last sentence is so unclear that even I'm not sure what it means. If this isn't apparent from the file, explain it much better! Wouldn't it be simpler if you just used numbering that started with 1 at the first codon?
Also, add the actual file so we can see if it makes sense.}
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:H1_prefs}
Amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/WSN/1933~\citep{doud2016accurate}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/WSN/1933 is in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:H3_prefs}
Amino acid preferences measured by the deep mutational scanning of the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/Perth/2009 is in in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:avg_prefs}
Average of the amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/Wilson Smith/1933~\citep{doud2016accurate} and the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the homologous sites between H1 HA and H3 HA. 
Conversion from this numbering scheme to sequential numbering in A/Wilson Smith/1933 or A/Perth/2009 is found in \ref{suppfile:WSN_Perth_map}. }
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:required_seqs}
This file contains the representative sequences of major human equine influenza lineages added to the alignment in \ref{suppfile:alignment_high}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_high}
The HA sequences used for the full HA tree in \ref{fig:empirical_trees}. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
\jdbcomment{This legend is not very good. If the file just has the alignment, then why are we talking about RAxML in its legend?
Also, I think referring to \ref{fig:empirical_trees} to refer to the tree topology is not very clear, since that figure shows lots of stuff beyond just one tree.}
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_lowH1}
The HA sequences using for the ``low divergence from H1" tree in \ref{fig:compete}. 
All sequences have $\ge 59\%$ amino-acid identity to the H1 reference sequence A/Wilson Smith/1933. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_lowH3}
The HA sequences using for the ``low divergence from H3" tree in \ref{fig:compete}. 
All sequences have $\ge 59\%$ amino-acid identity to the H3 reference sequence A/Perth/2009. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_intermediateH1}
The HA sequences using for the ``intermediate divergence from H1" tree in \ref{fig:compete}. 
All sequences have $\ge 46\%$ amino-acid identity to the H1 reference sequence A/Wilson Smith/1933. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_intermediateH3}
The HA sequences using for the ``intermediate divergence from H3" tree in \ref{fig:compete}. 
All sequences have $\ge 46\%$ amino-acid identity to the H3 reference sequence A/Perth/2009. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_GY94}
The GY94 model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_GY94+Gr}
The GY94+$\Gamma\omega$ model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_ExpCMdoud}
The ExpCM model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive} and the H1 HA amino-acid preference set in \ref{suppfile:H1_prefs}. 
}
\end{suppfile}
\clearpage

>>>>

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figures/empirical_trees.pdf}}
\caption{\label{fig:empirical_trees}
\textbf{Effect of site-specific amino-acid preferences and $\Gamma\omega$ rate variation on HA branch length estimation.} 
The branch lengths of the HA tree are optimized using the indicated ExpCM or GY94 model. 
The amino-acid preferences defining the model (ExpCM) or implied by the model (GY94) are shown as logo plots for 15 sites in HA; the full set of experimentally measured amino-acid preferences are in \ref{suppfig:prefs_doud}, \ref{suppfig:prefs_lee}, and \ref{suppfig:prefs_average}. 
The ExpCM's use amino-acid preferences measured in deep mutational scanning of an H1 HA~\citep{doud2016accurate}, an H3 HA~\citep{lee2018deep}, or the average of the measurements for these two HAs.
Circle denotes the H1 clade and triangle denotes the H3 clade.
\jdbcomment{Add a line to this legend about how trees are rooted, such as: ``Each tree is separately midpoint rooted''---or if that is not what was done, then add something that indicates how it was done. If it isn't what was done, can you remind me of rationale for not doing that?}
}
\end{figure}

@@

\begin{supptable}[t!]
\caption{\label{supptab:fit_params}
{\bf Total branch length for the trees in \ref{fig:empirical_trees}.}
The branch lengths are in average codon substitutions per site. 
} 
     \begin{tabular}{ccc}
        \hline
         Model & Parameters \\ \hline\hline
         GY94 &  {\shortstack{$\kappa = 3.17$, $\omega = 0.10$, \\$\phi_{1,A} = 0.32$, $\phi_{1,C} = 0.13$, $\phi_{1,G} = 0.28$, \\$\phi_{2,A} = 0.38$, $\phi_{2,C} = 0.18$,  $\phi_{2G} = 0.19$,\\ $\phi_{3, A} = 0.35$, $\phi_{3,C} = 0.19$, $\phi_{3,G} = 0.21$}}\\ \hline
         GY94 + $\Gamma\omega$ & {\shortstack{$\alpha_\omega = 0.51$, $\beta_\omega = 3.92$, $\kappa = 3.49$, \\$\phi_{1,A} = 0.32$, $\phi_{1,C} = 0.14$, $\phi_{1,G} = 0.28$, \\$\phi_{2,A} = 0.38$, $\phi_{2,C} = 0.18$, $\phi_{2,G} = 0.20$, \\$\phi_{3,A} = 0.355$, $\phi_{3,C} = 0.19$, $\phi_{3,G} = 0.21$}}\\ \hline
         ExpCM(H1) &  {\shortstack{$\beta = 1.22$, $\kappa = 3.24$, $\omega = 0.13$,\\ $\phi_A = 0.37$, $\phi_C = 0.18$, $\phi_G = 0.23$}}\\ \hline
      \end{tabular}
\end{supptable}
\clearpage

@@

\begin{suppfile}
\caption{
\label{suppfile:WSN_Perth_map}
List of alignable sites between H1 HA and H3 HA. 
This file provides a conversion between the numbering scheme we use in the paper (sequential numbering of just the alignable sites) to sequential numbering of the H1 HA reference sequence A/Wilson Smith/1933 and the H3 HA reference sequence A/Perth/2009. 
The sequential numbering for the HA strains starts with ``site 1" as the second codon (first non-methionine codon). 
\jdbcomment{The last sentence is so unclear that even I'm not sure what it means. If this isn't apparent from the file, explain it much better! Wouldn't it be simpler if you just used numbering that started with 1 at the first codon?
Also, add the actual file so we can see if it makes sense.}
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:H1_prefs}
Amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/WSN/1933~\citep{doud2016accurate}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/WSN/1933 is in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:H3_prefs}
Amino acid preferences measured by the deep mutational scanning of the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/Perth/2009 is in in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:alignment_high}
The HA sequences used for the full HA tree in \ref{fig:empirical_trees}. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
\jdbcomment{This legend is not very good. If the file just has the alignment, then why are we talking about RAxML in its legend?
Also, I think referring to \ref{fig:empirical_trees} to refer to the tree topology is not very clear, since that figure shows lots of stuff beyond just one tree.}
}
\end{suppfile}
